---
title: "EDA5-shuting"
author: "Shuting Li"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(plotly)
library(TSdist)
library(lubridate)
library(tibble)
library(factoextra)
library(rbenchmark)
library(autoencoder)
```

# import data
```{r}
actual_daily <- read_csv("~/Desktop/BU-2021fall/partner project/fidelity-github/11.14备份/lst/data/actual_daily.csv")
actual_daily$date <- as.character(actual_daily$date)

execution_daily <- read.csv("~/Desktop/BU-2021fall/partner project/fidelity-github/11.14备份/lst/data/execution_daily.csv")
```

# compare execution & actual credit

## loop for 320 warehouses to compute execution & actual credit distance
```{r}
actual_daily$date <-as.Date(actual_daily$date)
execution_daily$date <-as.Date(execution_daily$date)

## Define function to compute distance in a time period
exe_actual_distance <- function(begin,end,take_log=FALSE){
  dis_ac_exe <- rep(NA,320)
  acct_ave <- rep(NA,320)
  exec_ave <- rep(NA,320)
  actual_0 <- actual_daily[actual_daily$date >=begin & actual_daily$date<= end,]
  exe_0 <- execution_daily[execution_daily$date >=begin & execution_daily$date<= end,] 
  minact<-min(actual_0[actual_0$sum>0,]$sum)
  minexc<-min(exe_0[exe_0$sum>0,]$sum)
  for (i in 1:320){
    if(!take_log){
      actual <- actual_0[actual_0$NEWID==i,5] 
      exe <- exe_0[exe_0$NEWID==i,5]
      acct_ave[i]<-mean(unlist(actual))
      exec_ave[i]<-mean(unlist(exe))
      dis_ac_exe[i] <- DTWDistance(actual,exe, window.type="sakoechiba",window.size=1)
     }
    else{
      actual <- actual_0[actual_0$NEWID==i,5] 
      exe <- exe_0[exe_0$NEWID==i,5]
      logactual<-log(actual+minact)
      logexectut<-log(exe+minexc)  
      acct_ave[i]<-mean(unlist(logactual))
      exec_ave[i]<-mean(unlist(logexectut))
      dis_ac_exe[i] <- DTWDistance(logactual,logexectut, window.type="sakoechiba", window.size=1)
  }
  }
return(data.frame(actual=acct_ave, executed=exec_ave,distance=dis_ac_exe))
}


## add NEWID to output table
dis_ac_exe_final2 <-exe_actual_distance("2021-03-29","2021-07-04")
dis_ac_exe_final2 <- data.frame(NEWID=c(1:320),dis_ac_exe_final2)

dis_ac_exe_final_log <-exe_actual_distance("2021-03-29","2021-07-04",TRUE)
dis_ac_exe_final_log <- data.frame(NEWID=c(1:320),dis_ac_exe_final_log)

## make plot
par(mfrow=c(1,2))
plot(log(dis_ac_exe_final2$actual), log(dis_ac_exe_final2$distance), xlab="log_actual", ylab="log_distance")
plot(log(dis_ac_exe_final2$executed), log(dis_ac_exe_final2$distance), xlab="log_execution", ylab="log_distance")

ggplot(dis_ac_exe_final2)+geom_point()+
  aes(x=(executed), y=(actual), color=log(distance))+geom_abline(slope=1, intercept=0)+geom_quantile()+scale_x_log10()+scale_y_log10()

ggplot(dis_ac_exe_final2)+geom_point()+
  aes(x=(executed/actual), y=(distance))+geom_abline(slope=1, intercept=0)+geom_quantile()+scale_x_log10()+scale_y_log10()

```
log make less influence of gap between exe & act.


## Define Best K

```{r}
library(factoextra)
library(NbClust)
cls <- log(dis_ac_exe_final2[,-1]+1)
cls_scale <- scale(cls)

fviz_nbclust(cls,kmeans, method = "wss") + geom_vline(xintercept = 4, linetype = 2)  + labs(subtitle = "Elbow method")

fviz_nbclust(cls, kmeans, method = "silhouette") + labs(subtitle = "Silhouette method")

set.seed(123)
fviz_nbclust(cls, kmeans, nstart = 25,  method = "gap_stat", nboot = 50)+
  labs(subtitle = "Gap statistic method")

## NbClust
NbClust(cls, diss=NULL, distance = "euclidean", min.nc=2, max.nc=10, 
            method = "kmeans", index = "all")
NbClust(cls, diss=NULL, distance = "euclidean", min.nc=2, max.nc=15, 
            method = "complete", index = "all")
#According to the majority rule, the best number of clusters is  4
```
The plot above represents the variance within the clusters. It decreases as k increases, but it can be seen a bend at k = 4. This bend indicates that additional clusters beyond the fourth have little value. We’ll classify the observations into 4 clusters.

Elbow method: 4 clusters solution suggested
Silhouette method: 2 clusters solution suggested
Gap statistic method: 6 clusters solution suggested


## cluster warehouses based on distance between execution and actual credit
```{r}
## cluster
library(pheatmap)
set.seed(1)
pheatmap(log(dis_ac_exe_final2[,-1]+1))
fitres <- pheatmap(log(dis_ac_exe_final2[,-1]+1),kmeans_k = 4)
fitres$kmeans$cluster 

## add clustering info to original dataset
df_ac_exe_cluster2 <- cbind(dis_ac_exe_final2, cluster = fitres$kmeans$cluster)

```


## Validation
```{r}
exe_actual_validate_one_warehouse <- function(begin,end,ID1){
  ac_validate <- actual_daily[actual_daily$NEWID %in% c(ID1),]
  ac_validate <- ac_validate[ac_validate$date >=begin & ac_validate$date<= end,]
  exe_validate <- execution_daily[execution_daily$NEWID %in% c(ID1),]
  exe_validate <- exe_validate[exe_validate$date >=begin & exe_validate$date<= end,]
  datatemp<- data.frame(date= exe_validate$date, 
                        actual=ac_validate$sum, execute=exe_validate$sum)
  acplot <- ggplot(reshape2::melt(datatemp, id="date")) + 
    geom_line()+aes(x=date,y=value,color=variable)+
    ylab("credit")+ggtitle(paste("warehouse:",ID1))
  return(acplot)
}

## randomly pick one warehouse to check series pattern
date.start="2021-03-29"
date.end="2021-07-04"
gridExtra::grid.arrange(exe_actual_validate_one_warehouse(date.start,date.end,288),
                        exe_actual_validate_one_warehouse(date.start,date.end,157),
                        exe_actual_validate_one_warehouse(date.start,date.end,273),
                        ncol=3)
```

```{r}
# library("plot3D")
# scatter3D(log(dis_ac_exe_final2$actual+1), log(dis_ac_exe_final2$executed+1), log(dis_ac_exe_final2$distance+1))
```

```{r}
## validation update 2021.12.1
exe_actual_validate <- function(begin,end,ID1,ID2,ID3,ID4,ID5){
  ac_validate <- actual_daily[actual_daily$NEWID %in% c(ID1,ID2,ID3,ID4,ID5),]
  ac_validate <- ac_validate[ac_validate$date >=begin & ac_validate$date<= end,]
  exe_validate <- execution_daily[execution_daily$NEWID %in% c(ID1,ID2,ID3,ID4,ID5),]
  exe_validate <- exe_validate[exe_validate$date >=begin & exe_validate$date<= end,]
  acplot <- ggplot(ac_validate) + 
    aes(x=date,y=sum,color=factor(NEWID))+
    geom_point()+
    geom_smooth(method = "lm", se=F,color="black")+
    geom_line(alpha=0.5)+
    ylab("actual_credit")
  exeplot <- ggplot(exe_validate) +
    aes(x=date,y=sum,color=factor(NEWID))+
    geom_point()+
    geom_smooth(method = "lm", se=F,color="black")+
    geom_line()+
    ylab("execution_credit")
  return(list(acplot,exeplot))
}

exe_actual_validate("2020-11-29","2021-01-24",1,2,4,5,10)

```

